A Survey of Exoskeleton Control (Draft 1.1)

### 2015-09-19 paper.tex
more updates

@@

@@

Bagnellian behavior modeling.

@@

@@

@@

@@

Different types of actuation can be combined: springs and electric motors,
for example. This combination can also be viewed as combining energy
storage and actuation. Slow pneumatics and fast electric motors~\cite{Morimoto,UCLA}.
Mini/Macro~\cite{Stanford}.

@@

@@

stance leg requires a large torque and
relatively low bandwidth during the support period. However,
the swing leg undergoes a large motion, but it only supports
its own weight during the swing period. Therefore, the swing
leg requires a relatively small torque. Thus, a dual-mode
approach is very useful for an exoskeleton robot control.
In this paper, a dual-mode control method using a hydraulic
system is proposed,~\cite{IEEE07222598}

rise from squat, chair.
push large objects out of the way (standing full body push
or seated using legs to push)

@@

\begin{verbatim}
 rigid
  gears
   planetary
   other
  harmonic drive
  ball screw
  cosine attachment
  Sarcos Primus Humanoid knee - to reduce moment arm variation with angle
 hydraulic/pneumatic circuit
 variable
  belts
  cones
  fluid
  adjustable moment arm
   linear slider
   XY slider for actuation and ``joint'' on foot.
 series elastic actuation
\end{verbatim}

@@

Heat exhaust. Heatsink cooling, Force air colling, Water cooling.
Will housing and structure get too hot and fry operator?

@@

Variants of nonlinear feedback control such as feedback linearization or
sliding mode control are largely ignored. Once the decision to use feedback control
based on a set of observable quantities and with particular outputs is made,
one can try out the various linear and nonlinear feedback control paradigms
to see what works well.

@@

@@

@@

@@

Similar to counterbalancing or physical gravity compensation, except done
with exoskeleton actuators to cancel exoskeleton weight (but not inertia).
A simple formulation is:
\begin{equation}
\vtau = \mJ^{\tr} (m\vg)
\end{equation}
where $\vtau$ are the exoskeleton joint torques, 
$\mJ$ is a Jacobian matrix, $m$ is the weight of the exoskeleton,
and $\vg$ is the gravity vector. 
Operator relies on mechanical backdrivability to drive
exokeleton to do task.

@@

@@

The goal here is to make the exoskeleton imitate a desired linear dynamic system.
In the case where there is no force sensing between the operator and the
exoskeleton, the exoskeleton controller maps from exoskeleton configuration
and velocity to desired joint/actuator/synergy forces.
In this case the actuators need to generate forces and torques rather than
positions or angles, or linear or angular velocities.

The major drawback of impedance/ admittance control
is that a predefined trajectory or a predefined point must be
assigned as a desired value.[06990981] NOT REALLY TRUE, SEE INVERSE DYNAMICS.

Although impedance control are extensively utilized in rehabilitation
robotics [18], limited number of studies has been focused on
this method for power augmentation [19, 20].
\begin{verbatim}
[18] R. Riener, L. Lunenburger, S. Jezernik, M. Anderschitz, G. Colombo,
and V. Dietz, "Patient-cooperative strategies for robot-aided treadmill
training: first experimental results," Neural Systems and Rehabilitation
Engineering, IEEE Transactions on, vol. 13, pp. 380-394, 2005.
[19] B.-K. Lee, H.-D. Lee, J.-y. Lee, K. Shin, J.-S. Han, and C.-S. Han,
"Development of dynamic model-based controller for upper limb
exoskeleton robot," in Robotics and Automation (ICRA), 2012 IEEE
International Conference on, 2012, pp. 3173-3178.
[20] W. Yu, J. Rosen, and X. Li, "PID admittance control for an upper limb
exoskeleton," in American Control Conference (ACC), 2011, 2011, pp.
1124-1129.
\end{verbatim}

\subsection{Operator-Exoskeleton Force Control/Admittance Control}

The goal here is to use force sensing between the operator and the
exoskeleton to ultimately generate exoskeleton velocities or angular velocities,
imitating a desired linear dynamic system.
This type of control is often referred to as admittance control as the
exoskeleton maps contact forces into joint motion.
Inverse dynamics control (see below)
can be used to improve this type of control performance.

@@

@@

The exoskeleton moves as to create as little force as possible between the
operator and the exoskeleton. The actuators in this case may be force or
position sources, but ultimately operator-exoskeleton forces are mapped
to exoskeleton velocities or angular velocities.

This is how the Sarcos XOS/WEAR exoskeleton works.

\begin{equation}
\dot{\vq} = \mJ^{-1} (\mK_1 \vf)
\end{equation}
\begin{equation}
\vtau = \mK_2 \dot{\vq} + \gc
\end{equation}
where $\dot{\vq}$ are the exoskeleton joint velocities,
$\mJ$ is a Jacobian matrix, $\mK_1$ is a gain matrix,
$\vf$ is the operator-exoskeleton force, 
$\vtau$ is the exoskeleton joint torques, $\mK_2$ is a gain matrix,
and $\gc$ is gravity compensation.
[06990981]

Inverting a Jacobian matrix is problematic when the Jacobian matrices are
near singular, which is common in human gait and manipulation.
Using a fixed or gain scheduled gain matrix may make more sense.
\begin{equation}
\dot{\vq} = \mK_3 \vf
\end{equation}

@@

The exoskeleton moves as to provide a scale version of exoskeleton-world
contact forces, or some other (usually simple) mapping. 
The actuators in this case may be force or
position sources, but ultimately operator-exoskeleton forces are mapped
to exoskeleton velocities or angular velocities to control the
operator-exoskeleton forces to desired values.

@@

\subsection{Inverse dynamics control}

The goal here is to map from desired motion (in this case desired
accelerations) to exoskeleton actuator commands (and desired operator
and world contact forces).
Inverse dynamics control is a generalization of active gravity compensation to include
inertial, Coriolis, and centripetal forces (forces due to acceleration and
velocity).
Inverse dynamics can be added to open loop behavior execution, operator force
feedback control, impedance control, and admittance control.

A key question is where does desired acceleration come from?
\begin{verbatim}
  Operator-Exoskeleton Force Control/Admittance Control  
  Receding Horizon Control of optimal trajectory of 
    idealized LIPM+flywheel model (Atkeson)
  intent prediction (see above)
  neuromuscular model (Geyer, Wang, Van de Panne)
  hybrid: Optimal control + neuromuscular model (Atkeson NSF Proposal)
    - NMM provides bias through changes to optimization criterion
    - NMM provides hard constraints
    - NMM replaces components of model-based optimization scheme
\end{verbatim}

Friction, actuator dynamics, exo-world contact forces.

@@

\subsection{Sensitive Amplicfication Control}

BLEEX:
Sensitivity amplification control (SAC) was firstly
developed by Kazerooni [14]. In this method, the interaction
force from human are regarded as disturbance. The control
system has high sensitivity to disturbances and thus has fast
motion response to the interaction forces. The main drawback
of SAC is unwanted response to any disturbance regardless of
being intentional or non-intentional disturbance. In addition,
the highly accurate dynamic modelling is required in this
method.
\begin{verbatim}
H. Kazerooni, J.-L. Racine, L. Huang, and R. Steger, "On the control of
the berkeley lower extremity exoskeleton (BLEEX)," in Robotics and
Automation, 2005. ICRA 2005. Proceedings of the 2005 IEEE
International Conference on, 2005, pp. 4353-4360.
\end{verbatim}

@@

stance: virtual joint torque control method Kazerooni et al.(2005)
When contact location between the wearer and the
exoskeleton is not fixed and difficult to estimate, this method
has been shown to be an effective method to generate the
locomotion for an exoskeleton robot.[3],[4],[21]
H. Kazerooni , Z. Racine , L. Huang and R. Steger ”On the control of
the Berkeley lower extremity exoskeleton (BLEEX),” Proc. IEEE Int.
Conf. Robot. Autom., pp. 4364-4371, 2005.
Racine Jean-Louis Charles ”Control of a Lower Extrmity Exoskeleton
for Human Performance Amplification,” University of California,2003.
582
Xiuxia Yang, Hongchao Zhao, Yi Zhang, Xiaowei Liu ”Carrying
Lower Extreme Exoskeleton Rapid Terminal Sliding-Mode Robust
Control,” Journal of computers, vol.9, No.1,202-208. 2014.

to reduce
sudden changes(command jump) at the phase transitions, a
smoothing method [5],[19] is introduced and a pre-transition
method is used to solve the swing delay due to the internal
pressure(approx 5 bar).
H. Kazerooni , Ryan Steger, Lihua Huang ”Hybrid Control of the
Berkeley Lower Extremity Exoskeleton (BLEEX),” Int. Journal of
Robotics Research, pp. 561-573, 2006.
oonbum Bae, Kyoungchul Kong, Masayoshi Tomizuka ”Gait Phase-
Based Control for a Rotary Series Elastic Actuator Assisting the knee

@@

\subsection{Symbiosis and Autonomy, Handling Errors, and Superhuman Reflexes}

@@

>>>>

Here is an example of a robot seeing and moving faster than a human, and
consistantly beating a human opponent in rock-paper-scissors.
\url{https://www.youtube.com/watch?v=Qb5UIPeFClM&feature=player_embedded}

@@

Predictive models of human motion, a major topic of robot learning
from observation/demonstration/imitiation and computer animation,
will be the subject of a future white
paper~\cite{IEEE06913830,Bagnell}

@@

@@

muscle moves -> surface electrodes come off (especially for large fast
movements, the worst time).
perspration changes measurement. needs to be calibrated for each application.

{\bf Foot contact and force/torque sensing:}
\begin{verbatim}
[8] K. Kong and M. Tomizuka, “A Gait Monitoring System Based on Air
Pressure Sensors Embedded in a Shoe,” IEEE/ASME Transactions on
Mechatronics, vol. 14, pp. 358–370, 2009.
\end{verbatim}

@@

\subsection{References}

\begin{verbatim}
H.Kawamoto and Y.Sankai, “Power Assist Method Based on Phase
Sequence Driven by Interaction between Human and Robot Suit,”
Proceedings of the IEEE International Workshop on Robot and Human
Interactive Communication, pp. 491–496, 2004.
\end{verbatim}

@@

Maxon dc flat brushless motor EC45 and servo controller
Elmo. The servomotor emulates the kinematic profile of every
joint and sends a digital signal to the Elmo to form feedbacks.
In order to enhance the impedance capability, harmonic trans-
mission drivers are implemented (model SHD-17-100-2SH
for joints 1 and 2, model SHD-14-100-2SH for joints 3 and 4,
and CSF-32-50-2A-GR for joint 5).~\cite{IEEE07128705}

@@

@@

\subsection{Hybrid actuation}

Different types of actuation can be combined: springs and electric motors,
for example. This combination can also be viewed as combining energy
storage and actuation. Slow pneumatics and fast electric motors~\cite{Morimoto,UCLA}.
Mini/Macro~\cite{Stanford}. In this view, the human operator is just
one more set of actuators to be added to the combined system.

\begin{verbatim}
parallel combinations - weak fast actuator needs to hit joint limit and then be strong.
 (leads to chattering) or just lock in position to exert big force
series combination - strong actuator needs to clutch out of circuit to move fast
\end{verbatim}

@@

Stance leg requires a large torque and
relatively low bandwidth during the support period. However,
the swing leg undergoes a large motion, but it only supports
its own weight during the swing period. Therefore, the swing
leg requires a relatively small torque. Thus, a dual-mode
approach is very useful for an exoskeleton robot control.
In this paper, a dual-mode control method using a hydraulic
system is proposed,~\cite{IEEE07222598}

Rise from squat, chair.
push large objects out of the way (standing full body push
or seated using legs to push)

@@

\begin{verbatim}
 rigid
  gears
   planetary
   other
  harmonic drive
  ball screw
  cosine attachment
  Sarcos Primus Humanoid knee - to reduce moment arm variation with angle
 Bowden cable
 tendon drives (can go slack)
 hydraulic/pneumatic circuit
 variable
  belts
  cones
  fluid
  adjustable moment arm
   linear slider
   XY slider for actuation and ``joint'' on foot.
 series elastic actuation
\end{verbatim}

@@

Heat exhaust. Heatsink cooling, Force air cooling, Water cooling.
Will housing and structure get too hot and fry operator?

@@

Variants of nonlinear feedback control such as feedback linearization or
sliding mode control are largely ignored. Once the decision to use feedback control
based on a set of observable quantities and with particular outputs is made,
one can try out the various linear and nonlinear feedback control paradigms
to see what works well.
Variants of function approximation methods such as lookup tables, fuzzy logic,
sigmoidal neural networks, radial basis functions, and locally weighted regression
are also largely ignored. Once the decision to use a function approximator
and what the inputs and outputs has been made, one can try out the various
approaches to see what works well.
Same for optimization methods.
Same for constraint enforcement (such as avoiding self-collisions)
in either optimization or feedback control (barrier
Lyapunov functions (BLF)~\cite{IEEE06911561}).

@@

The dynamics of a revolute exoskeleton are:
\begin{equation}
\mM(\vq) \ddot{\vq} + \mC(\vq,\dot{\vq}) + \mG(\vq) + \mF(\vq,\dot{\vq})
+ \mJ_w^{\tr}(\vq) \vf_w = \vtau + \mJ_o^{\tr}(\vq) \vf_o
\end{equation}
where $\vq$ are the joint angles and position and velocity of the ``root'' of
the exoskeleton, $\mM$ is the inertia matrix, $\mC()$ are the Coriolis and
centripetal forces, $\mG()$ are the gravitational forces, $\mF$ are the
friction forces, $\mJ_w^{\tr}(\vq) \vf_w$ (Jacobian matrix multiplied with the
contact forces)
are the exoskeleton-world
contact forces 
expressed as exoskeleton joint torques, 
$\vtau$ are the exoskeleton
joint torques,
and $\mJ_o^{\tr}(\vq) \vf_o$
are the operator-exoskelton forces expressed
as exoskeleton joint torques,

To understand these equations better, it is useful to have all Jacobians equal
to the identity matrix. This means the operator is directly applying torques
at the exoskeleton joints, and so is the world. In this case:
\begin{equation}
\mM(\vq) \ddot{\vq} + \mC(\vq,\dot{\vq}) + \mG(\vq) + \mF(\vq,\dot{\vq})
+ \vtau_w = \vtau + \vtau_o
\end{equation}

Note that we are ignoring the operator dynamics and operator muscle and tissue
forces resulting in operator joint torques. We will add these in later (???).
We are also ignoring actuator dynamics, which we can handle by augmenting 
the state vector.

@@

$\vtau$ is generated by passive devices such as springs or dampers.

@@

$\vtau = \vtau_{ff}(t)$ where $\vtau_{ff}$ is a time
dependent feedforward torque vector.

@@

Similar to counterbalancing or physical gravity compensation, except done
with exoskeleton actuators to cancel exoskeleton weight (but not inertia).
A simple formulation is:
\begin{equation}
\vtau = \mJ^{\tr} (m\vg)
\end{equation}
where $\vtau$ are the exoskeleton joint torques, 
$\mJ$ is a Jacobian matrix, $m$ is the weight of the exoskeleton,
and $\vg$ is the gravity vector. 
The operator relies on mechanical backdrivability to drive
the exokeleton to do tasks.

@@

$\vtau = \hat{\mG}(\hat{\vq})$ 
are the estimated gravitational torques at the estimated exoskeleton configuration.

\subsection{Friction compensation}

$\vtau = \hat{\mF}(\hat{\vq},\hat{\dot{\vq}})$ 
are the estimated friction torques at the estimated exoskeleton configuration
and velocity.
Note that these may be load dependent, and may require additional state to
correctly handle stiction and hysteresis effects.

\subsection{Online inverse dynamics control}

The goal here is to map from desired motion (in this case desired
accelerations) to exoskeleton actuator commands (and desired operator
and world contact forces).
The exoskeleton actuator torque 
attempts to match the torques caused by inertial, Coriolis, centripetal,
gravitational, frictional, and exoskeleton-world contact forces.
It relies on estimated dynamic models and their parameters.
If there are actuator dynamics or other dynamics, we can include those
dynamics as well.
\begin{equation}
\vtau = \vtau_{ID}
\hat{\mM}(\hat{\vq}) \ddot{\vq}_d
+ \hat{\mC}(\hat{\vq},\hat{\dot{\vq}})
+ \hat{\mG}(\hat{\vq})
+ \hat{\mF}(\hat{\vq},\hat{\dot{\vq}})
+ \hat{\mJ}_w^{\tr}(\hat{\vq}) \hat{\vf}_w
\end{equation}
If the models and estimated state (position and velocity) are perfect, we get
perfect cancellation of most of the dynamics:
\begin{equation}
\mM(\vq) (\ddot{\vq} - \ddot{\vq}_d) = \mJ_o^{\tr}(\vq) \vf_o
\end{equation}
If $\ddot{\vq}_d$ is set to zero, the operator directly drives the acceleration
of the exoskeleton with an effective inertia of $(\mM^{-1}(\vq) \mJ_o^{\tr}(\vq))^{-1}$
\begin{equation}
\ddot{\vq} = \mM^{-1}(\vq) \mJ_o^{\tr}(\vq) \vf_o
\end{equation}
$\ddot{\vq}_d$ can be used to drive the exoskeleton along trajectories, and
the operator can directly modify the trajectory.

Simplifying Equation X by assuming the operator directly applies torques at the
exoskeleton joints we get:
\begin{equation}
\mM (\vq) \ddot{\vq} = \vtau_o
\end{equation}
The operator sees the true inertia of the exoskeleton, but no gravitational
loads and no other dynamics (Coriolis, centripetal, or frictional).
Note that we cannot change the apparent inertia of the exoskeleton without
some form of acceleration or force feedback.

Online inverse dynamics control is a generalization of active gravity compensation to include
inertial, Coriolis, and centripetal forces (forces due to acceleration and
velocity).
If frictional forces are modeled they can also be canceled.
Online inverse dynamics can be added to open loop behavior execution
and many other types of control such as operator force
feedback control, impedance control, and admittance control.
Computed torque control and feedback linearization are forms of online
inverse dynamics.

A key question is where does desired acceleration come from?
\begin{verbatim}
  Operator force control.
  Receding Horizon Control of optimal trajectory of 
    idealized LIPM+flywheel model (Atkeson)
  intent prediction (see above)
  neuromuscular model (Geyer, Wang, Van de Panne)
  hybrid: Optimal control + neuromuscular model (Atkeson NSF Proposal)
    - NMM provides bias through changes to optimization criterion
    - NMM provides hard constraints
    - NMM replaces components of model-based optimization scheme
\end{verbatim}

@@

The goal here is to make the exoskeleton imitate a desired linear dynamic system
from the point of view of the operator.
A future white paper will discuss how to make the exoskeleton 
imitate a desired linear dynamic system 
from the point of view of external perturbations.
In the case where there is no force sensing between the operator and the
exoskeleton, the exoskeleton controller maps from exoskeleton configuration
and velocity to desired joint/actuator/synergy forces.
In this case the actuators need to generate forces and torques rather than
positions or angles, or linear or angular velocities.

Here we make the exoskeleton appear as a locally linear impedance in joint
coordinates to the operator.
We can also make the exoskeleton appear as a locally linear impedance to the
operator in some other coordinate system, such as Cartesian coordinates.
A different goal is to make the exoskeleton appear as a locally linear impedance to
the outside world in some coordinate system.

To make the exoskeleton appear as a locally linear impedance in joint
coordinates to the operator, we add position and velocity feedback to $\vtau$:
\begin{equation}
\vtau = \vtau_{ID} - \mK_{\vq} ( \hat{\vq} - \vq_d ) - \mK_{\dot{\vq}} \hat{\dot{\vq}}
\end{equation}
so if the dynamic models and state estimation are perfect we get this impedance:
\begin{equation}
\mM \ddot{\vq} - \mK_{\vq} ( \vq - \vq_d ) - \mK_{\dot{\vq}} \dot{\vq} = \vtau_o
\end{equation}
and the operator sees the desirecd impedance.

Although impedance control is extensively utilized in rehabilitation
robotics [18], limited number of studies has been focused on
this method for power augmentation [19, 20].
\begin{verbatim}
[18] R. Riener, L. Lunenburger, S. Jezernik, M. Anderschitz, G. Colombo,
and V. Dietz, "Patient-cooperative strategies for robot-aided treadmill
training: first experimental results," Neural Systems and Rehabilitation
Engineering, IEEE Transactions on, vol. 13, pp. 380-394, 2005.
[19] B.-K. Lee, H.-D. Lee, J.-y. Lee, K. Shin, J.-S. Han, and C.-S. Han,
"Development of dynamic model-based controller for upper limb
exoskeleton robot," in Robotics and Automation (ICRA), 2012 IEEE
International Conference on, 2012, pp. 3173-3178.
[20] W. Yu, J. Rosen, and X. Li, "PID admittance control for an upper limb
exoskeleton," in American Control Conference (ACC), 2011, 2011, pp.
1124-1129.
\end{verbatim}

\subsection{Force feedback and get out of the way control (inverse dynamics version)}

The goal here is to use force sensing between the operator and the
exoskeleton to ultimately generate exoskeleton velocities or angular velocities,
imitating a desired linear dynamic system.
This type of control is often referred to as admittance control as the
exoskeleton maps contact forces into joint motion.
Inverse dynamics control
can be used to improve this type of control performance.

@@

We can implement force control 
if measurements of the contact forces 
the exoskeleton are available.
The exoskeleton moves as to create as little force as possible 
assuming the actuators are torque sources.
\begin{equation}
\vtau = \vtau_{ID} - \mK_{\vf} \mJ^{\tr} ( \hat{\vf} - \vf_d )
\end{equation}
$\mK_{\vf}$ is a gain matrix, $\mJ$ is the appropriate Jacobian matrix for
force $\vf$
Note that this does not require inverting the Jacobian matrix.
We will see that the position control version does invert the Jacobian matrix.

Applying this to force between the operator and the exoskeleton, and assuming
that the operator directly applies exoskeleton joint torques,
\begin{equation}
\vtau = \vtau_{ID} - \mK_{\vtau_o} ( \hat{\vtau}_o - {\vtau_o}_d )
\end{equation}
So if our dynamic models and state estimation are perfect, we get:
\begin{equation}
\ddot{\vq} = - \mM^{-1} (\vq) \mK_{\vtau_o} ( \hat{\vtau}_o - {\vtau_o}_d )
\end{equation}

One can add force damping terms, which sometimes help:
\begin{equation}
\vtau = \vtau_{ID} - \mK_{\vtau_o} ( \hat{\vtau}_o - {\vtau_o}_d )
- \mK_{\dot{\vtau}_o} \hat{\dot{\vtau}}_o
\end{equation}

Note that impedance control can also be implemented using force control
by setting:
\begin{equation}
{\vtau_o}_d = \mK_{\vq}( \hat{\vq} - \vq_d ) + \mK_{\dot{\vq}} \hat{\dot{\vq}}
\end{equation}
in addition to directly including the stiffness and damping terms in
equation YYY.

\subsection{Operator-Exoskeleton Force Control using position controlled actuation}

At this point our dynamics formulation is no longer appropriate, since we
are assuming the actuators are now position rather than torque sources.

@@

The exoskeleton moves as to create as little force as possible between the
operator and the exoskeleton. The actuators in this case may be position sources or
forces sources with a high servo gains,
but ultimately operator-exoskeleton forces are mapped
to exoskeleton velocities or angular velocities.

\begin{equation}
\dot{\vq}_d = \mJ^{-1} (\mK_1 \vf)
\end{equation}
where $\dot{\vq}_d$ are the commanded exoskeleton joint velocities,
$\mJ$ is an appropriate Jacobian matrix, $\mK_1$ is a gain matrix,
$\vf$ is the force vector to be controlled.
[06990981]

Inverting a Jacobian matrix is problematic when the Jacobian matrices are
near singular.
Using a fixed or gain scheduled gain matrix may make more sense.
\begin{equation}
\dot{\vq}_d = \mK_3 \vf
\end{equation}
Simplifying Equation Z by assuming the forces to be controlled
are directly applied to the exoskeleton joints, and thus the Jacobian matrix
is the identity matrix, gives us a similar equation:
\begin{equation}
\dot{\vq}_d = \mK_1 \vtau_o
\end{equation}
and we avoid inverting a Jacobian and problems with singularities.

@@

We introduce a desired force to allow for more complex force control.
\begin{equation}
\dot{\vq}_d = \mJ^{-1} (\mK_1 (\vf - \vf_d))
\end{equation}

The exoskeleton could move as to provide a scaled version of exoskeleton-world
contact forces, or some other (usually simple) mapping. 
\begin{equation}
\vf_d = \alpha \mJ_{ow} \vf_w
\end{equation}

@@

@@

\subsection{Sensitivity Amplification Control}

Sensitivity Amplification Control (SAC) is a dynamic cancellation technique (a.k.a.
inverse dynamics, computed torque, or feedback linearization).

BLEEX:
Sensitivity amplification control (SAC) was
developed by Kazerooni [14]. In this method, the interaction
forces with the human are regarded as disturbances. The control
system has high sensitivity to disturbances and thus has fast
motion response to the interaction forces. The main drawback
of SAC is unwanted responses to any disturbance regardless of
being intentional or non-intentional disturbance. In addition,
highly accurate dynamic modelling is required in this
method.
\begin{verbatim}
H. Kazerooni, J.-L. Racine, L. Huang, and R. Steger, "On the control of
the berkeley lower extremity exoskeleton (BLEEX)," in Robotics and
Automation, 2005. ICRA 2005. Proceedings of the 2005 IEEE
International Conference on, 2005, pp. 4353-4360.
\end{verbatim}

@@

stance: virtual joint torque control method Kazerooni et al.(2005)
When contact location between the wearer and the
exoskeleton is not fixed and difficult to estimate, this method
has been shown to be an effective method to generate the
locomotion for an exoskeleton robot.[3],[4],[21]
\begin{verbatim}
H. Kazerooni , Z. Racine , L. Huang and R. Steger ”On the control of
the Berkeley lower extremity exoskeleton (BLEEX),” Proc. IEEE Int.
Conf. Robot. Autom., pp. 4364-4371, 2005.
Racine Jean-Louis Charles ”Control of a Lower Extrmity Exoskeleton
for Human Performance Amplification,” University of California,2003.
582
Xiuxia Yang, Hongchao Zhao, Yi Zhang, Xiaowei Liu ”Carrying
Lower Extreme Exoskeleton Rapid Terminal Sliding-Mode Robust
Control,” Journal of computers, vol.9, No.1,202-208. 2014.
\end{verbatim}

to reduce
sudden changes(command jump) at the phase transitions, a
smoothing method [5],[19] is introduced and a pre-transition
method is used to solve the swing delay due to the internal
pressure(approx 5 bar).
\begin{verbatim}
H. Kazerooni , Ryan Steger, Lihua Huang ”Hybrid Control of the
Berkeley Lower Extremity Exoskeleton (BLEEX),” Int. Journal of
Robotics Research, pp. 561-573, 2006.
oonbum Bae, Kyoungchul Kong, Masayoshi Tomizuka ”Gait Phase-
Based Control for a Rotary Series Elastic Actuator Assisting the knee
\end{verbatim}

@@

\section{Control Architectures}

A generic hierarchical control architecture~\cite{IEEE06907051}:

{\bf Lowest level Control:} present interface to actuators and
high bandwidth feedback control, degrees of freedom,
or synergies of either:
\begin{itemize}
\item
Position control.
\item
Pure force or torque control.
\item
Impedance control: specify desired position (for spring), and generate
stiffness, damping, and potentially a modified mass/iinertia/moment of inertia.
Modifying inertia is very difficult.
Or (almost equivalently),
specify a desired force and change the force according to position, velocity,
and acceleration.
\end{itemize} 
One important synergy is coordinated control of the center of mass location and
velocity for balance.
PID or fancier feedback control laws are implemented at this level.
Gravity, friction, actuator dynamics, and other feedforward disturbance
compensation is implemented at this level, as well as any inverse dynamics
or other types of feedforward reference tracking control.
Disturbance observers for model errors and contact/wind forces
are implemented at this level.
Typically this is done by a high rate global servo~\cite{Atlas-robot}
or local limb or joint-level servos~\cite{Sarcos-robot}.

{\bf ``Phoneme'' Control:} mid-level controllers
for relatively constant phases of behavior.
Often involves position or velocity targets (goto or
stay-at X), trajectory references, ...
Example from walking: left-stance, double-support-1, right-stance, double-support-2, ...;
Example from running: left-foot-down, flight-1, right-foot-down, flight-2, ...;
Example from horizontal jumping: hjump-pushoff, hjump-flight, hjump-prepare-landing, hjump-impact, hjump-balance, ...;
Finer resolution may be needed. For example a walking stance period may
be broken up into heel-strike, first-half, second-half, pushoff, and toe-off.
The first and second halves of stance are divided by when the center of mass
passes over the ankle.
Local inverse kinematics (finding small joint motions for small target changes)
is performed at this level.

{\bf ``Word'' Control:} mid-level controllers for sequences of ``phoneme'' behavioral
units.
Often involves finite state machines, timers, if then condition testing, gain
scheduling, ...
Examples: walk-to-location( x=7.8m, y=2.3m ), run-to-location( speed=3.7m/s ),
horizontal-jump( height= 2m, length= 5m ).
Global inverse kinematics (selection of solution branch) is performed at this level.

{\bf Highest level control} aka {\bf Behavior Selection:} 
Answer the question: ``What do I do next?''
Select behaviors and behavioral parameters (targets,
speeds, durations).
Often involves finite state machines. timers, if then condition testing, ...

{\bf Error handling:} Errors must be detected and handled by behavior switching
at all of these levels.

An alternative behavioral architecture does not have explicit levels but instead
has a ``soup'' of behaviors that compete or cooperate for control.

Another alternative behavioral architecture does not have explicit separate behaviors,
but one big function (produced by deep learning?).

\section{Symbiosis and Autonomy, Handling Errors, and Superhuman Reflexes}

@@

S.Viteckova, P.Kutilek, and M.Jirina, “Wearable Lower Limb Robotics:
A Review,” Biocybernetics and Biomedical Engineering, vol. 33, pp.
96–105, 2013.
